Supplementary Material for: A Rough Guide to Pre-processing High-Frequency Animal Tracking Data

### 2020-11-23 docs/supplementary_material.tex
add supplement tex

]{article}

@@

\usepackage{lineno}
\setmainfont{Times New Roman}

@@

\title{Supplementary Material for: A Rough Guide to Pre-processing High-Frequency Animal Tracking Data}
\author{Pratik R. Gupte \and Christine E. Beardsworth \and Orr Spiegel \and Emmanuel Lourie \and Allert Bijleveld}
\date{2020-11-16}

@@

{
\setcounter{tocdepth}{2}
\tableofcontents
}
\hypertarget{processing-egyptian-fruit-bat-tracks}{%
\section{Processing Egyptian fruit bat tracks}\label{processing-egyptian-fruit-bat-tracks}}

@@

Install the required \texttt{R} libraries that are available on CRAN.

@@

\begin{figure}
\centering
\includegraphics{figures/fig_bat_patches.png}
\caption{A visual examination of plots of the bats' residence patches and linear approximations of paths between them showed that though all three bats roosted at the same site, they used distinct areas of the study site over the three nights \textbf{(a)}.
Bats tended to be resident near fruit trees, which are their main food source, travelling repeatedly between previously visited areas \textbf{(b, c)}.
However, bats also appeared to spend some time at locations where no fruit trees were recorded, prompting questions about their use of other food sources \textbf{(b, c)}.
When bats did occur close together, their residence patches barely overlapped, and their paths to and from the broad area of co-occurrence were not similar \textbf{(c)}.
Constructing residence patches for multiple individuals over multiple activity periods suggests interesting dynamics of within- and between-individual overlap \textbf{(b, c)}.}
\end{figure}

@@

\hypertarget{refs}{}
\begin{cslreferences}
\leavevmode\hypertarget{ref-bracis2018}{}%
Bracis, Chloe, Keith L. Bildstein, and Thomas Mueller. 2018. ``Revisitation Analysis Uncovers Spatio-Temporal Patterns in Animal Movement Data.'' \emph{Ecography} 41 (11): 1801--11. \url{https://doi.org/10.1111/ecog.03618}.

@@

>>>>

]{scrartcl}

@@

\usepackage{lineno}
\setmainfont{Times New Roman}
\setsansfont{Arial}

@@

\title{Supplementary Material for: A Rough Guide to Pre-processing High-Frequency Animal Tracking Data}
\author{Pratik R. Gupte \and Christine E. Beardsworth \and Orr Spiegel \and Emmanuel Lourie \and Allert Bijleveld}
\date{2020-11-23}

@@

{
\setcounter{tocdepth}{2}
\tableofcontents
}
\hypertarget{processing-calibration-data}{%
\section{Processing calibration data}\label{processing-calibration-data}}

Here we show how the residence patch method (Barraquand and Benhamou \protect\hyperlink{ref-barraquand2008}{2008}; Bijleveld et al. \protect\hyperlink{ref-bijleveld2016}{2016}; Oudman et al. \protect\hyperlink{ref-oudman2018}{2018}) accurately estimates the duration of known stops in a track collected as part of a calibration exercise in the Wadden Sea.

First we prepare the libraries we need. Libraries can be installed from CRAN if necessary.

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\# load libs}
\KeywordTok{library}\NormalTok{(data.table)}
\KeywordTok{library}\NormalTok{(atlastools)}
\KeywordTok{library}\NormalTok{(ggplot2)}
\KeywordTok{library}\NormalTok{(patchwork)}

\hypertarget{access-data-and-preliminary-visualisation}{%
\subsection{Access data and preliminary visualisation}\label{access-data-and-preliminary-visualisation}}

First we access the data from a local file using the \texttt{data.table} package (Dowle and Srinivasan \protect\hyperlink{ref-dowle2020}{2020}).
We then visualise the raw data.

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\# read and plot example data}
\NormalTok{data <{-}}\StringTok{ }\KeywordTok{fread}\NormalTok{(}\StringTok{"data/atlas1060\_allTrials\_annotated.csv"}\NormalTok{)}
\NormalTok{data\_raw <{-}}\StringTok{ }\KeywordTok{copy}\NormalTok{(data)}
\end{Highlighting}
\end{Shaded}

\includegraphics{figures/fig_calibration_raw.png}

\hypertarget{filter-by-bounding-box}{%
\subsection{Filter by bounding box}\label{filter-by-bounding-box}}

We first save a copy of the data, so that we can plot the raw data with the cleaned data plotted over it for comparison.

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\# make a copy using the data.table copy function}
\NormalTok{data\_unproc <{-}}\StringTok{ }\KeywordTok{copy}\NormalTok{(data)}
\end{Highlighting}
\end{Shaded}

We then filter by a bounding box in order to remove the point outlier to the far south east of the main track. We use the \texttt{atl\_filter\_bounds} functions using the \texttt{x\_range} argument, to which we pass the limit in the UTM 31N coordinate reference system.
This limit is used to exclude all points with an X coordinate \textless{} 645,000.

We then plot the result of filtering, with the excluded point in black, and the points that are retained in green.

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\# remove inside must be set to falses}
\NormalTok{data <{-}}\StringTok{ }\KeywordTok{atl\_filter\_bounds}\NormalTok{(}\DataTypeTok{data =}\NormalTok{ data, }
                          \DataTypeTok{x =} \StringTok{"x"}\NormalTok{, }\DataTypeTok{y =} \StringTok{"y"}\NormalTok{, }
                          \DataTypeTok{x\_range =} \KeywordTok{c}\NormalTok{(}\DecValTok{645000}\NormalTok{, }\KeywordTok{max}\NormalTok{(data}\OperatorTok{$}\NormalTok{x)), }
                          \DataTypeTok{remove\_inside =} \OtherTok{FALSE}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\hypertarget{filter-trajectories}{%
\subsection{Filter trajectories}\label{filter-trajectories}}

\hypertarget{handle-time}{%
\subsubsection{Handle time}\label{handle-time}}

Time in ATLAS tracking is counted in milliseconds and is represented by a 64-bit integer (type \texttt{long}), which is not natively supported in R; it will instead be converted to a \texttt{numeric}, or \texttt{double}.

This is not what is intended, but it works. The \texttt{bit64} package can help handle 64-bit integers if you want to keep to intended type.

A further issue is that 64-bit integers (whether represented as \texttt{bit64} or \texttt{double}) do not yield meaninful results when you try to convert them to a date-time object, such as of the class \texttt{POSIXct}.

This is because \texttt{as.POSIXct} fails when trying to work with 64-bit integers (it cannot interpret this type), and returns a date many thousands of years in the future (approx. 52,000 CE) if the time column is converted to \texttt{numeric}.

There are two possible solutions. The parsimonious one is to convert the 64-bit number to a 32-bit short integer (dividing by 1000), or to use the \texttt{nanotime} package.

The conversion method loses an imperceptible amount of precision. The \texttt{nanotime} requires installing another package. The first method is shown here.

In the spirit of not destroying data, we create a second lower-case column called \texttt{time}.

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\# divide by 1000, convert to integer, then convert to POSIXct}
\NormalTok{data[, time }\OperatorTok{:}\ErrorTok{=}\StringTok{ }\KeywordTok{as.integer}\NormalTok{(TIME }\OperatorTok{/}\StringTok{ }\DecValTok{1000}\NormalTok{)]}
\end{Highlighting}
\end{Shaded}

\hypertarget{add-speed-and-turning-angle}{%
\subsubsection{Add speed and turning angle}\label{add-speed-and-turning-angle}}

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\# add incoming and outgoing speed}
\NormalTok{data[, }\StringTok{\textasciigrave{}}\DataTypeTok{:=}\StringTok{\textasciigrave{}}\NormalTok{ (}\DataTypeTok{speed\_in =} \KeywordTok{atl\_get\_speed}\NormalTok{(data, }
                                      \DataTypeTok{x =} \StringTok{"x"}\NormalTok{, }
                                      \DataTypeTok{y =} \StringTok{"y"}\NormalTok{, }
                                      \DataTypeTok{time =} \StringTok{"time"}\NormalTok{),}
             \DataTypeTok{speed\_out =} \KeywordTok{atl\_get\_speed}\NormalTok{(data, }\DataTypeTok{type =} \StringTok{"out"}\NormalTok{))]}

\CommentTok{\# add turning angle}
\NormalTok{data[, angle }\OperatorTok{:}\ErrorTok{=}\StringTok{ }\KeywordTok{atl\_turning\_angle}\NormalTok{(}\DataTypeTok{data =}\NormalTok{ data)]}
\end{Highlighting}
\end{Shaded}

\hypertarget{get-95th-percentile-of-speed-and-angle}{%
\subsubsection{Get 95th percentile of speed and angle}\label{get-95th-percentile-of-speed-and-angle}}

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\# use sapply}
\NormalTok{speed\_angle\_thresholds <{-}}\StringTok{ }
\StringTok{  }\KeywordTok{sapply}\NormalTok{(data[, }\KeywordTok{list}\NormalTok{(speed\_in, speed\_out, angle)], }
\NormalTok{       quantile, }\DataTypeTok{probs =} \FloatTok{0.9}\NormalTok{, }\DataTypeTok{na.rm =}\NormalTok{ T)}
\end{Highlighting}
\end{Shaded}

\hypertarget{filter-on-speed}{%
\subsubsection{Filter on speed}\label{filter-on-speed}}

Here we use a speed threshold of 15 m/s, the fastest known boat speed.
We then plot the data with the extreme speeds shown in grey, and the positions retained shown in green.

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\# make a copy}
\NormalTok{data\_unproc <{-}}\StringTok{ }\KeywordTok{copy}\NormalTok{(data)}

\CommentTok{\# remove speed outliers}
\NormalTok{data <{-}}\StringTok{ }\KeywordTok{atl\_filter\_covariates}\NormalTok{(}\DataTypeTok{data =}\NormalTok{ data,}
            \DataTypeTok{filters =} \KeywordTok{c}\NormalTok{(}\StringTok{"(speed\_in < 15 \& speed\_out < 15)"}\NormalTok{))}

\CommentTok{\# recalculate speed and angle}
\NormalTok{data[, }\StringTok{\textasciigrave{}}\DataTypeTok{:=}\StringTok{\textasciigrave{}}\NormalTok{ (}\DataTypeTok{speed\_in =} \KeywordTok{atl\_get\_speed}\NormalTok{(data, }
                                      \DataTypeTok{x =} \StringTok{"x"}\NormalTok{, }
                                      \DataTypeTok{y =} \StringTok{"y"}\NormalTok{, }
                                      \DataTypeTok{time =} \StringTok{"time"}\NormalTok{),}
             \DataTypeTok{speed\_out =} \KeywordTok{atl\_get\_speed}\NormalTok{(data, }\DataTypeTok{type =} \StringTok{"out"}\NormalTok{))]}

\CommentTok{\# add turning angle}
\NormalTok{data[, angle }\OperatorTok{:}\ErrorTok{=}\StringTok{ }\KeywordTok{atl\_turning\_angle}\NormalTok{(}\DataTypeTok{data =}\NormalTok{ data)]}
\end{Highlighting}
\end{Shaded}

\includegraphics{figures/fig_speed_outlier.png}

\hypertarget{smoothing-the-trajectory}{%
\subsection{Smoothing the trajectory}\label{smoothing-the-trajectory}}

We then apply a median smooth over a moving window (\(K\) = 5).
This function modifies in place, and does not need to be assigned to a new variable.
We create a copy of the data before applying the smooth so that we can compare the data before and after smoothin.

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\# apply a 5 point median smooth, first make a copy}
\NormalTok{data\_unproc <{-}}\StringTok{ }\KeywordTok{copy}\NormalTok{(data)}

\CommentTok{\# now apply the smooth}
\KeywordTok{atl\_median\_smooth}\NormalTok{(}\DataTypeTok{data =}\NormalTok{ data,}
                  \DataTypeTok{x =} \StringTok{"x"}\NormalTok{, }\DataTypeTok{y =} \StringTok{"y"}\NormalTok{, }\DataTypeTok{time =} \StringTok{"time"}\NormalTok{,}
                  \DataTypeTok{moving\_window =} \DecValTok{5}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\includegraphics{figures/fig_calib_median_smooth.png}

\hypertarget{thinning-the-data}{%
\subsection{Thinning the data}\label{thinning-the-data}}

Next we thin the data to demonstrate thinning by median smoothing.
Following this, we plot the median smooth and thinning by aggregation.

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\# save a copy}
\NormalTok{data\_unproc <{-}}\StringTok{ }\KeywordTok{copy}\NormalTok{(data)}

\CommentTok{\# remove columns we don\textquotesingle{}t need}
\NormalTok{data <{-}}\StringTok{ }\NormalTok{data[, }\KeywordTok{setdiff}\NormalTok{(}\KeywordTok{colnames}\NormalTok{(data), }
                       \KeywordTok{c}\NormalTok{(}\StringTok{"tID"}\NormalTok{, }\StringTok{"Timestamp"}\NormalTok{, }\StringTok{"id"}\NormalTok{, }\StringTok{"TIME"}\NormalTok{, }\StringTok{"UTCtime"}\NormalTok{)), }
\NormalTok{             with =}\StringTok{ }\OtherTok{FALSE}\NormalTok{]}

\CommentTok{\# thin to a 30s interval}
\NormalTok{data\_thin <{-}}\StringTok{ }\KeywordTok{atl\_thin\_data}\NormalTok{(}\DataTypeTok{data =}\NormalTok{ data,}
                           \DataTypeTok{interval =} \DecValTok{30}\NormalTok{,}
                           \DataTypeTok{method =} \StringTok{"aggregate"}\NormalTok{,}
                           \DataTypeTok{id\_columns =} \StringTok{"TAG"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\includegraphics{figures/fig_calib_smooth_thin.png}

\hypertarget{residence-patches}{%
\subsection{Residence patches}\label{residence-patches}}

\hypertarget{get-waypoint-centroids}{%
\subsubsection{Get waypoint centroids}\label{get-waypoint-centroids}}

We subset the annotated calibration data to select the waypoints and the positions around them which are supposed to be the locations of known stops. Since each stop was supposed to be 5 minutes long, there are multiple points in each known stop.

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{library}\NormalTok{(stringi)}
\NormalTok{data\_res <{-}}\StringTok{ }\NormalTok{data\_unproc[}\KeywordTok{stri\_detect}\NormalTok{(tID, }\DataTypeTok{regex =} \StringTok{"(WP)"}\NormalTok{)]}
\end{Highlighting}
\end{Shaded}

From this data, we get the centroid of known stops, and determine the time difference between the first and last point within 50 metres, and within 10 minutes of the waypoint positions' median time.

Essentially, this means that the maximum duration of a stop can be 20 minutes, and stops above this duration are not expected.

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\# get centroid}
\NormalTok{data\_res\_summary <{-}}\StringTok{ }\NormalTok{data\_res[, }\KeywordTok{list}\NormalTok{(}\DataTypeTok{x\_median =} \KeywordTok{round}\NormalTok{(}\KeywordTok{median}\NormalTok{(x), }\DataTypeTok{digits =} \DecValTok{{-}2}\NormalTok{),}
                                    \DataTypeTok{y\_median =} \KeywordTok{round}\NormalTok{(}\KeywordTok{median}\NormalTok{(y), }\DataTypeTok{digits =} \DecValTok{{-}2}\NormalTok{)),}
\NormalTok{                                    t\_median =}\StringTok{ }\KeywordTok{median}\NormalTok{(time)}\ErrorTok{)}\NormalTok{,}
\NormalTok{                             by =}\StringTok{ "tID"}\NormalTok{]}

\CommentTok{\# now get times 10 mins before and after}
\NormalTok{data\_res\_summary[, }\StringTok{\textasciigrave{}}\DataTypeTok{:=}\StringTok{\textasciigrave{}}\NormalTok{(}\DataTypeTok{t\_min =}\NormalTok{ t\_median }\OperatorTok{{-}}\StringTok{ }\NormalTok{(}\DecValTok{10} \OperatorTok{*}\StringTok{ }\DecValTok{60}\NormalTok{),}
                        \DataTypeTok{t\_max =}\NormalTok{ t\_median }\OperatorTok{+}\StringTok{ }\NormalTok{(}\DecValTok{10} \OperatorTok{*}\StringTok{ }\DecValTok{60}\NormalTok{))]}

\CommentTok{\# make a list of positions 10min before and after}
\NormalTok{wp\_data <{-}}\StringTok{ }\KeywordTok{mapply}\NormalTok{(}\ControlFlowTok{function}\NormalTok{(l, u, mx, my) \{}
\NormalTok{  tmp\_data <{-}}\StringTok{ }\NormalTok{data\_unproc[}\KeywordTok{inrange}\NormalTok{(time, l, u)]}
\NormalTok{  tmp\_data[, distance }\OperatorTok{:}\ErrorTok{=}\StringTok{ }\KeywordTok{sqrt}\NormalTok{((mx }\OperatorTok{{-}}\StringTok{ }\NormalTok{x)}\OperatorTok{\^{}}\DecValTok{2} \OperatorTok{+}\StringTok{ }\NormalTok{(my }\OperatorTok{{-}}\StringTok{ }\NormalTok{y)}\OperatorTok{\^{}}\DecValTok{2}\NormalTok{)]}

  \CommentTok{\# keep within 50}
\NormalTok{  tmp\_data <{-}}\StringTok{ }\NormalTok{tmp\_data[distance }\OperatorTok{<=}\StringTok{ }\DecValTok{50}\NormalTok{, ]}

  \CommentTok{\# get duration}
  \KeywordTok{return}\NormalTok{(}\KeywordTok{diff}\NormalTok{(}\KeywordTok{range}\NormalTok{(tmp\_data}\OperatorTok{$}\NormalTok{time)))}

\NormalTok{\}, data\_res\_summary}\OperatorTok{$}\NormalTok{t\_min, data\_res\_summary}\OperatorTok{$}\NormalTok{t\_max,}
\NormalTok{   data\_res\_summary}\OperatorTok{$}\NormalTok{x\_median, data\_res\_summary}\OperatorTok{$}\NormalTok{y\_median,}
\DataTypeTok{SIMPLIFY =} \OtherTok{TRUE}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\hypertarget{prepare-data}{%
\subsubsection{Prepare data}\label{prepare-data}}

An indicator of individual residence at or near a position can be useful when attempting to identify residence patches. Positions can be filtered on a metric such as residence time (Bracis, Bildstein, and Mueller \protect\hyperlink{ref-bracis2018}{2018}).

\hypertarget{calculate-residence-time}{%
\subsubsection{Calculate residence time}\label{calculate-residence-time}}

First we calculate the residence time with a radius of 50 metres.
For this, we need a dataframe with coordinates, the timestamp, and the animal id.
We save this data to file for later use.

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\# load recurse}
\KeywordTok{library}\NormalTok{(recurse)}
\end{Highlighting}
\end{Shaded}

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\# get 4 column data}
\NormalTok{data\_for\_patch <{-}}\StringTok{ }\NormalTok{data\_thin[, }\KeywordTok{list}\NormalTok{(x, y, time, TAG)]}

\CommentTok{\# get recurse data for a 10m radius}
\NormalTok{recurse\_stats <{-}}\StringTok{ }\KeywordTok{getRecursions}\NormalTok{(data\_for\_patch,}
                               \DataTypeTok{radius =} \DecValTok{50}\NormalTok{, }\DataTypeTok{timeunits =} \StringTok{"mins"}\NormalTok{)}

\CommentTok{\# assign to recurse data}
\NormalTok{data\_for\_patch[, res\_time }\OperatorTok{:}\ErrorTok{=}\StringTok{ }\NormalTok{recurse\_stats}\OperatorTok{$}\NormalTok{residenceTime]}

\CommentTok{\# save recurse data}
\KeywordTok{fwrite}\NormalTok{(data\_for\_patch, }\DataTypeTok{file =} \StringTok{"data/data\_calib\_for\_patch.csv"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\hypertarget{run-residence-patch-method}{%
\subsubsection{Run residence patch method}\label{run-residence-patch-method}}

We subset data with a residence time \textgreater{} 5 minutes in order to construct residence patches.
From this subset, we construct residence patches using the parameters: \texttt{buffer\_radius} = 5 metres, \texttt{lim\_spat\_indep} = 50 metres, \texttt{lim\_time\_indep} = 5 minutes, and \texttt{min\_fixes} = 3.

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\# assign id as tag}
\NormalTok{data\_for\_patch[, id }\OperatorTok{:}\ErrorTok{=}\StringTok{ }\KeywordTok{as.character}\NormalTok{(TAG)]}

\CommentTok{\# on known residence points}
\NormalTok{patch\_res\_known <{-}}\StringTok{ }\KeywordTok{atl\_res\_patch}\NormalTok{(data\_for\_patch[res\_time }\OperatorTok{>=}\StringTok{ }\DecValTok{5}\NormalTok{, ], }
                                \DataTypeTok{buffer\_radius =} \DecValTok{5}\NormalTok{,}
                                \DataTypeTok{lim\_spat\_indep =} \DecValTok{50}\NormalTok{,}
                                \DataTypeTok{lim\_time\_indep =} \DecValTok{5}\NormalTok{,}
                                \DataTypeTok{min\_fixes =} \DecValTok{3}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\hypertarget{get-spatial-and-summary-objects}{%
\subsubsection{Get spatial and summary objects}\label{get-spatial-and-summary-objects}}

We get spatial and summary ouput of the residence patch method using the \texttt{atl\_patch\_summary} function using the options \texttt{which\_data} = ``spatial'' and \texttt{which\_data} = "summary.
We use a buffer radius here of 20 metres for the spatial buffer, despite using a buffer radius of 5 metres earlier, simply because it is easier to visualise in the output figure.

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\# for the known and unkniwn patches}
\NormalTok{patch\_sf\_data <{-}}\StringTok{ }\KeywordTok{atl\_patch\_summary}\NormalTok{(patch\_res\_known, }
                                   \DataTypeTok{which\_data =} \StringTok{"spatial"}\NormalTok{,}
                                   \DataTypeTok{buffer\_radius =} \DecValTok{20}\NormalTok{)}

\CommentTok{\# assign crs}
\NormalTok{sf}\OperatorTok{::}\KeywordTok{st\_crs}\NormalTok{(patch\_sf\_data) <{-}}\StringTok{ }\DecValTok{32631}

\CommentTok{\# get summary data}
\NormalTok{patch\_summary\_data <{-}}\StringTok{ }\KeywordTok{atl\_patch\_summary}\NormalTok{(patch\_res\_known, }
                                        \DataTypeTok{which\_data =} \StringTok{"summary"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\hypertarget{prepare-to-plot-data}{%
\subsubsection{Prepare to plot data}\label{prepare-to-plot-data}}

We read in the island's shapefile to plot it as a background for the residence patch figure.

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\# read griend}
\NormalTok{griend <{-}}\StringTok{ }\NormalTok{sf}\OperatorTok{::}\KeywordTok{st\_read}\NormalTok{(}\StringTok{"data/griend\_polygon/griend\_polygon.shp"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\hypertarget{compare-patch-metrics}{%
\subsection{Compare patch metrics}\label{compare-patch-metrics}}

We then merge the annoated, known stop data with the calculated patch duration.
We filter this data to exclude one exceedingly long outlier of about an hour (WP080), which how

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\# get known patch summary}
\NormalTok{data\_res <{-}}\StringTok{ }\NormalTok{data\_unproc[stringi}\OperatorTok{::}\KeywordTok{stri\_detect}\NormalTok{(tID, }\DataTypeTok{regex =} \StringTok{"(WP)"}\NormalTok{), ]}

\CommentTok{\# get waypoint summary}
\NormalTok{patch\_summary\_real <{-}}\StringTok{ }\NormalTok{data\_res[, }\KeywordTok{list}\NormalTok{(}\DataTypeTok{nfixes\_real =}\NormalTok{ .N,}
                                      \DataTypeTok{x\_median =} \KeywordTok{round}\NormalTok{(}\KeywordTok{median}\NormalTok{(x), }\DataTypeTok{digits =} \DecValTok{{-}2}\NormalTok{),}
                                      \DataTypeTok{y\_median =} \KeywordTok{round}\NormalTok{(}\KeywordTok{median}\NormalTok{(y), }\DataTypeTok{digits =} \DecValTok{{-}2}\NormalTok{)), }
\NormalTok{                               by =}\StringTok{ "tID"}\NormalTok{]}

\CommentTok{\# add real duration}
\NormalTok{patch\_summary\_real[, duration\_real }\OperatorTok{:}\ErrorTok{=}\StringTok{ }\NormalTok{wp\_data]}

\CommentTok{\# round median coordinate for inferred patches}
\NormalTok{patch\_summary\_inferred <{-}}\StringTok{ }
\StringTok{  }\NormalTok{patch\_summary\_data[, }
                     \KeywordTok{c}\NormalTok{(}\StringTok{"x\_median"}\NormalTok{, }\StringTok{"y\_median"}\NormalTok{, }
                       \StringTok{"nfixes"}\NormalTok{, }\StringTok{"duration"}\NormalTok{, }\StringTok{"patch"}\NormalTok{)}
\NormalTok{                     ][, }\StringTok{\textasciigrave{}}\DataTypeTok{:=}\StringTok{\textasciigrave{}}\NormalTok{(}\DataTypeTok{x\_median =} \KeywordTok{round}\NormalTok{(x\_median, }\DataTypeTok{digits =} \DecValTok{{-}2}\NormalTok{),}
                              \DataTypeTok{y\_median =} \KeywordTok{round}\NormalTok{(y\_median, }\DataTypeTok{digits =} \DecValTok{{-}2}\NormalTok{))]}

\CommentTok{\# join with respatch summary}
\NormalTok{patch\_summary\_compare <{-}}\StringTok{ }
\StringTok{  }\KeywordTok{merge}\NormalTok{(patch\_summary\_real,}
\NormalTok{        patch\_summary\_inferred, }
        \DataTypeTok{on =} \KeywordTok{c}\NormalTok{(}\StringTok{"x\_median"}\NormalTok{, }\StringTok{"y\_median"}\NormalTok{),}
        \DataTypeTok{all.x =} \OtherTok{TRUE}\NormalTok{, }\DataTypeTok{all.y =} \OtherTok{TRUE}\NormalTok{)}

\CommentTok{\# drop nas}
\NormalTok{patch\_summary\_compare <{-}}\StringTok{ }\KeywordTok{na.omit}\NormalTok{(patch\_summary\_compare)}

\CommentTok{\# drop patch around WP080}
\NormalTok{patch\_summary\_compare <{-}}\StringTok{ }\NormalTok{patch\_summary\_compare[tID }\OperatorTok{!=}\StringTok{ "WP080"}\NormalTok{, ]}
\end{Highlighting}
\end{Shaded}

7 patches are identified where there are no waypoints, while 2 waypoints are not identified as patches. These waypoints consisted of 6 and 15 (WP098 and WP092) positions respectively, and were lost when the data were aggregated to 30 second intervals.

\hypertarget{linear-model-durations}{%
\subsubsection{Linear model durations}\label{linear-model-durations}}

We run a simple linear model.

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\# get linear model}
\NormalTok{model\_duration <{-}}\StringTok{ }\KeywordTok{lm}\NormalTok{(duration\_real }\OperatorTok{\textasciitilde{}}\StringTok{ }\NormalTok{duration,}
                     \DataTypeTok{data =}\NormalTok{ patch\_summary\_compare)}

\CommentTok{\# get R2}
\KeywordTok{summary}\NormalTok{(model\_duration)}

\CommentTok{\# write to file}
\KeywordTok{writeLines}\NormalTok{(}
  \DataTypeTok{text =} \KeywordTok{capture.output}\NormalTok{(}
    \KeywordTok{summary}\NormalTok{(model\_duration)}
\NormalTok{  ),}
  \DataTypeTok{con =} \StringTok{"data/model\_output\_residence\_patch.txt"}
\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\includegraphics{figures/fig_calib_lm_duration.png}

\hypertarget{processing-egyptian-fruit-bat-tracks}{%
\section{Processing Egyptian fruit bat tracks}\label{processing-egyptian-fruit-bat-tracks}}

@@

Install the required \texttt{R} libraries that are required from CRAN if not already installed.

@@

\begin{figure}
\centering
\includegraphics{figures/fig_08_bats.png}
\caption{A visual examination of plots of the bats' residence patches and linear approximations of paths between them showed that though all three bats roosted at the same site, they used distinct areas of the study site over the three nights \textbf{(a)}.
Bats tended to be resident near fruit trees, which are their main food source, travelling repeatedly between previously visited areas \textbf{(b, c)}.
However, bats also appeared to spend some time at locations where no fruit trees were recorded, prompting questions about their use of other food sources \textbf{(b, c)}.
When bats did occur close together, their residence patches barely overlapped, and their paths to and from the broad area of co-occurrence were not similar \textbf{(c)}.
Constructing residence patches for multiple individuals over multiple activity periods suggests interesting dynamics of within- and between-individual overlap \textbf{(b, c)}.}
\end{figure}

@@

\hypertarget{refs}{}
\begin{cslreferences}
\leavevmode\hypertarget{ref-barraquand2008}{}%
Barraquand, Frédéric, and Simon Benhamou. 2008. ``Animal Movements in Heterogeneous Landscapes: Identifying Profitable Places and Homogeneous Movement Bouts.'' \emph{Ecology} 89 (12): 3336--48. \url{https://doi.org/10.1890/08-0162.1}.

\leavevmode\hypertarget{ref-bijleveld2016}{}%
Bijleveld, Allert Imre, Robert B MacCurdy, Ying-Chi Chan, Emma Penning, Richard M. Gabrielson, John Cluderay, Erik L. Spaulding, et al. 2016. ``Understanding Spatial Distributions: Negative Density-Dependence in Prey Causes Predators to Trade-Off Prey Quantity with Quality.'' \emph{Proceedings of the Royal Society B: Biological Sciences} 283 (1828): 20151557. \url{https://doi.org/10.1098/rspb.2015.1557}.

\leavevmode\hypertarget{ref-bracis2018}{}%
Bracis, Chloe, Keith L. Bildstein, and Thomas Mueller. 2018. ``Revisitation Analysis Uncovers Spatio-Temporal Patterns in Animal Movement Data.'' \emph{Ecography} 41 (11): 1801--11. \url{https://doi.org/10.1111/ecog.03618}.

@@

\leavevmode\hypertarget{ref-oudman2018}{}%
Oudman, Thomas, Theunis Piersma, Mohamed V. Ahmedou Salem, Marieke E. Feis, Anne Dekinga, Sander Holthuijsen, Job ten Horn, Jan A. van Gils, and Allert I. Bijleveld. 2018. ``Resource Landscapes Explain Contrasting Patterns of Aggregation and Site Fidelity by Red Knots at Two Wintering Sites.'' \emph{Movement Ecology} 6 (1): 24--24. \url{https://doi.org/10.1186/s40462-018-0142-4}.
